---
title: "CAS MA 585 Final Project"
author: "Joel Choe Yee Hsien (U22167011)"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/Joel Choe/Desktop/CAS MA 585/Final Project")
library(tidyverse)
library(TSA)
library(tseries)
library(MASS)
library(forecast)
library(urca)
library(stats4)
library(sarima)
library(ggplot2)
library(dplyr)
library(lubridate)
library(astsa)
library(gridExtra)
library(anytime)
```


Load dataset
```{r a}
food_price_index <- read_csv('food_price_index.csv') %>% filter(Series_title_1 == 'Food')
data <- food_price_index[, c("Period", "Data_value")]
data <- data %>% filter(Period > 2000)
colnames(data) <- c('date', 'value')
data.start_time <- ymd(anydate(as.character(data$date[[1]]))) %>% print()
data.end_time <- ymd(anydate(as.character(data$date[[nrow(data)]]))) %>% print()
data.freq <- 12
```

Check for outliers

```{r b}
ggplot(data) +
  aes(x = "", y = value) +
  geom_boxplot() +
  theme_minimal()
```

Convert data to time series and plot data points

```{r c}
data_all <- ts(data$value, start = c(year(data.start_time), month(data.start_time)), end = c(year(data.end_time), month(data.end_time)), frequency = data.freq)
plot(data_all, main = "Monthly Food Price Index in New Zealand", xlab = "Year", ylab = "Food Price Index")
```

Basic Description of the data

```{r d}
min(data_all)
max(data_all)
mean(data_all)
```

Test for stationarity

```{r e}
kpss.test(data_all)
adf.test(data_all)
```

Classical Decomposition

```{r f}
autoplot(decompose(data_all))
```

ACF and PACF

```{r g}
ggtsdisplay(data_all)
```

Split dataset to generate train and test set

```{r h}
data_train <- head(data_all, round(0.8 * length(data_all)))
data_test <- tail(data_all, round(length(data_all) - round(0.8 * length(data_all))))
print(length(data_train))
print(length(data_test))
```

Finding a suitable SARIMA model

```{r i}
data_train_lag <- diff(data_train, lag = data.freq)
ggtsdisplay(data_train_lag, main  = "Lag-12 Seasonal Differencing", ylab  = "Food Price Index", xlab  = "Year")
```

```{r j}
data_train_diff_lag = diff(diff(data_train, lag = 2), lag = 12)
ggtsdisplay(data_train_diff_lag, main  = "Lag-12 + Diff-2 Seasonal Differencing", ylab  = "Food Price Index", xlab  = "Year")
```

```{r k}
kpss.test(data_train_diff_lag)
adf.test(data_train_diff_lag)
```

First ARMA Model

```{r l}
arma_model = Arima(data_train, order = c(3, 2, 1), seasonal = list(order = c(1, 1, 1), period = data.freq))
checkresiduals(arma_model)
tsdiag(arma_model)
qqnorm(arma_model$residuals)
qqline(arma_model$residuals)
AIC(arma_model)
BIC(arma_model)
summary(arma_model)
```

Fitting an ARMA Model

```{r m}
arma_model = Arima(data_train, order = c(3, 2, 3), seasonal = list(order = c(1, 1, 1), period = data.freq))
checkresiduals(arma_model)
tsdiag(arma_model)
qqnorm(arma_model$residuals)
qqline(arma_model$residuals)
AIC(arma_model)
BIC(arma_model)
summary(arma_model)
```

```{r n}
arma_model_pred <- forecast(arma_model, h = length(data_test))
plot(arma_model_pred)
lines(data_test)
```

```{r o}
accuracy(arma_model_pred, data_test)
```


Fitting an AR Model

```{r p}
ar_model = Arima(data_train, order = c(3, 2, 0), seasonal = list(order = c(1, 1, 0), period = data.freq))
checkresiduals(ar_model)
tsdiag(ar_model)
qqnorm(ar_model$residuals)
qqline(ar_model$residuals)
AIC(ar_model)
BIC(ar_model)
summary(ar_model)
```

```{r q}
ar_model_pred <- forecast(ar_model, h = length(data_test))
plot(ar_model_pred)
lines(data_test)
```

```{r s}
accuracy(ar_model_pred, data_test)
```

Fitting a MA model

```{r t}
ma_model = Arima(data_train, order = c(0, 2, 3), seasonal = list(order = c(0, 1, 1), period = data.freq))
checkresiduals(ma_model)
tsdiag(ma_model)
qqnorm(ma_model$residuals)
qqline(ma_model$residuals)
AIC(ma_model)
BIC(ma_model)
summary(ma_model)
```

```{r u}
ma_model_pred <- forecast(ma_model, h = length(data_test))
plot(ma_model_pred)
lines(data_test)
```

```{r v}
accuracy(ma_model_pred, data_test)
```

Brute Force Algo

AIC

```{r w}
p_list <- c()
q_list <- c()
P_list <- c()
Q_list <- c()
aic_score <- c()
minimum <- Inf
best <- c()
for(p in c(0, 1, 2, 3, 4)) {
  for(q in c(0, 1, 2, 3, 4)) {
    for(P in c(0, 1)) {
      for(Q in c(0, 1)) {
        arma_model <- Arima(data_train, order = c(p, 2, q), seasonal = list(order = c(P, 1, Q), period = data.freq))
        p_list <- c(p_list, p)
        q_list <- c(q_list, q)
        P_list <- c(P_list, P)
        Q_list <- c(Q_list, Q)
        aic_score <- c(aic_score, AIC(arma_model))
        if(AIC(arma_model) < minimum) {
          minimum <- AIC(arma_model)
          best <- c(p, q, P, Q)
          print(AIC(arma_model))
          print(best)
        }
      }
    }
  }
}
```

```{r x}
arma_model_aic <- Arima(data_train, order = c(2, 2, 3), seasonal = list(order = c(1, 1, 1), period = data.freq))
checkresiduals(arma_model_aic)
tsdiag(arma_model_aic)
qqnorm(arma_model_aic$residuals)
qqline(arma_model_aic$residuals)
AIC(arma_model_aic)
BIC(arma_model_aic)
summary(arma_model_aic)
```

```{r y}
arma_model_aic_pred <- forecast(arma_model_aic, h = length(data_test))
plot(arma_model_aic_pred, main = "Prediction of Food Price Index of SARIMA(2,2,3)(1,1,1)")
lines(data_test)
```

```{r ae}
accuracy(arma_model_aic_pred, data_test)
```

BIC

```{r z}
p_list <- c()
q_list <- c()
P_list <- c()
Q_list <- c()
bic_score <- c()
minimum <- Inf
best <- c()
for(p in c(0, 1, 2, 3, 4)) {
  for(q in c(0, 1, 2, 3, 4)) {
    for(P in c(0, 1)) {
      for(Q in c(0, 1)) {
        arma_model <- Arima(data_train, order = c(p, 2, q), seasonal = list(order = c(P, 1, Q), period = data.freq))
        p_list <- c(p_list, p)
        q_list <- c(q_list, q)
        P_list <- c(P_list, P)
        Q_list <- c(Q_list, Q)
        bic_score <- c(bic_score, BIC(arma_model))
        if(BIC(arma_model) < minimum) {
          minimum <- BIC(arma_model)
          best <- c(p, q, P, Q)
          print(BIC(arma_model))
          print(best)
        }
      }
    }
  }
}
```

```{r aa}
arma_model_bic <- Arima(data_train, order = c(0, 2, 1), seasonal = list(order = c(0, 1, 1), period = data.freq))
checkresiduals(arma_model_bic)
tsdiag(arma_model_bic)
qqnorm(arma_model_bic$residuals)
qqline(arma_model_bic$residuals)
AIC(arma_model_bic)
BIC(arma_model_bic)
summary(arma_model_bic)
```

```{r ab}
arma_model_bic_pred <- forecast(arma_model_bic, h = length(data_test))
plot(arma_model_bic_pred)
lines(data_test)
```

```{r af}
accuracy(arma_model_bic_pred, data_test)
```

Holt-Winters Forecast

```{r ac}
model_HW_add <- HoltWinters(data_train, seasonal = "additive")
pred_HW_add <- forecast(model_HW_add, h = length(data_test))
plot(pred_HW_add)
lines(data_test)
```

```{r ad}
accuracy(pred_HW_add, data_test) %>% print()
```

```{r ag}
model_HW_add_full <- HoltWinters(data_all, seasonal = "additive")
pred_HW_add_full <- forecast(model_HW_add_full, h = 16)
plot(pred_HW_add_full)
pred_HW_add_full
```


